---
title: "Lab3"
author: "Heather M"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


data <- readxl::read_xls('Lab3.xls')
library(ggplot2) 
library(dplyr)
library(tidyverse)
head(data)
```
# 1.Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find. 


```{r}
rename(data, polparty = 'Political party affiliation')
```
```


```{r}
Dem <- filter(data, data$'Political party affiliation' == 'Strong democrat')
Rep <- filter(data, data$'Political party affiliation' == 'Strong republican')
Indep <- filter(data, data$'Political party affiliation'  == 'Independent')
ggplot(Dem, aes(x=Dem$'Political party affiliation', y=Dem$'Highest year of school completed')) + geom_point()
```
```{r}
ggplot(Indep, aes(x=Indep$'Political party affiliation', y=Indep$'Highest year of school completed')) + geom_line()
```

```{r}
ggplot(data, aes(x='Political party affiliation',)) + geom_boxplot()
```
#2.Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.
For each of the investigation below, create at least one numerical summary and one appropriate visualization. Write a paragraph to describe your findings and the variables you used.
```{r}
Dem2 <- select(Dem, 'Political party affiliation','Highest year of school completed')
Demo <- Dem2 %>% top_frac(.85)
Dem3 <- head(Demo, n = 14)
Dem3
```
```{r}
ggplot(Dem3, aes(x='Political party affiliation', y='Highest year of school completed')) + geom_point()
```

```{r}
Rep2 <- select(Rep, 'Political party affiliation','Highest year of school completed')
Repu <- Rep2 %>% top_frac(.85)
Rep3 <- head(Repu, n = 14)
Rep3
```
```{r}
Indep2 <- select(Indep, 'Political party affiliation','Highest year of school completed')
Ind <- Indep2 %>% top_frac(.85)
Ind3 <- head(Ind, n = 14)
Ind3
```

Based the data I have in Dem3, Ind3, and Rep3, It looks like each party has a similar education year difference. The avgerage is around 12-13 which is around a high school diploma. That is fair to say most people have at least a high school diploma.  

#Does party affiliation affect the relationship you found in the last question?

#No, not that I have seen. All of the averages are very similar to one another. One affiliation that was different was that Independents were slightly lower average than the democrats and republicans

#How is marrital status related to the education of a respondent? What are possible reasons accounting for the patterns you see?
It doesn't really have a strong effect on education of a respondent. There are an array of people that come from all different marital statuses. There are a lot more married people who answered the survey than single individuals. 
#Explore two more interesting questions, and answer those questions using the GSS data.
 1. In the Democratic party what race partcipated the most?
```{r}
EduDem <- Dem$'Highest year of school completed'
RaceDem <- Dem$'Race of respondent'
ggplot(Dem, aes(x=RaceDem, )) + geom_bar()
```
#It looks like by a large amount white people were the biggest contributors to this survey
#2 In this data set who participated the most, Democrats, Republicans, or Indepedents?
```{r}
PolParty <- data$'Political party affiliation'
ggplot(data, aes(x=PolParty, dev.new(width =5,height=8))) + geom_bar()

```

It looks like strong democrats participated in the survey the most, according to the graph. 



